Examining the Multifaceted Impact of Family Size on Happiness, Health, Gender Dynamics, and Financial Well-being

### 2023-04-05 outputs/paper/paper.qmd
Add files via upload

---
title: "Examining the Multifaceted Impact of Family Size on Happiness, Health, Gender Dynamics, and Financial Well-being"
subtitle: ""
author: 
  - SHAOHAN CHANG
thanks: "Code and data are available at: https://github.com/lucas11333/Of_Wellbeing"
date: "`r Sys.time()`"
date-format: "D MMMM YYYY"
abstract: "This study analyzed the relationship between family size, measured by the number of children, and various aspects of family life, such as happiness, health, and financial satisfaction. The findings suggest that happiness levels are highest among childless individuals, while families with one child report the best health status. Additionally, higher financial satisfaction is associated with fewer children in a family. These insights matter as they provide a better understanding of the factors influencing family size and the potential consequences of family planning decisions on overall wellbeing."
format: pdf
number-sections: true
bibliography: references.bib
---

@@

The impact of the number of children in a family on various aspects of family life has been a topic of interest for researchers and policymakers alike [@a]. Family size and structure play a crucial role in shaping the lives of individuals and communities. Understanding the connections between the number of children and factors such as happiness, health, sex distribution, and financial satisfaction can help inform decisions and policies aimed at promoting family well being [@b]. This paper seeks to explore these relationships and shed light on the complex interplay between family size and various aspects of family life.

@@

In @tbl-var contains information on various personal and demographic characteristics of individuals, collected between the years 1972 and 2012. The data-set includes variables such as age, sex, number of babies, divorce status, self-reported health status, financial satisfaction and level of happiness. The data-set can be used to study the relationships between these variables and to gain insights into factors that may influence an individual's overall life satisfaction.

@@

In [@fig-one],which analysis of this data is to calculate the number of children with 0, 1, 2, 3, 4.through the graph showing the relationship between the number of children and happiness in the family, it is obvious that when people have no children, the largest number of 'very happy', 'pretty happy', and 'not too happy' are shown. Although The preliminary judgment is that the number of people who are happy without children is the largest. But among the data, the data collected without children is the most. In order to ensure the accuracy of the data, it is necessary to further analyze the proportion of the number of children in the family for the next part of the analysis.

@@

In [@fig-four], the analysis of the chart, it can be observed that when individuals have one child, there is a significantly higher proportion of participants who rate their health as 'good' or 'excellent' compared to other categories. In contrast, for participants who do not have any children, the proportion of those who rate their health as 'poor' is the highest, but no clear conclusion can be drawn from this finding. For individuals with three children, the feedback on their health status is evenly distributed among 'good', 'fair', and 'excellent', with a very small proportion of participants reporting 'poor' health. Overall, having one child is associated with the highest proportion of participants reporting 'excellent' health status.

@@

$$
\hat{Y}=\hat{\beta_0}+\hat{\beta_1}X_{finance-score}+\hat{\beta_2}X_{health-score} + \hat{\beta_3}X_{happy} + 
\hat{\beta_6}X_{babies-number} 
$$ {#eq-model}

@@

-   ${\hat{Y}}$ is predicted value of the dependent variable. This is the outcome we're trying to predict or estimate based on the independent variables.
-   ${\hat{\beta_0}}$ : The intercept term, which represents the expected value of ${\hat{Y}}$ when all independent variables are equal to 0.
-   ${\hat{\beta_1}}$ , ${\hat{\beta_2}}$ , ${\hat{\beta_3}}$ are the estimated coefficients for each independent variable. These coefficients indicate the average change in the dependent variable ${\hat{Y}}$ associated with a one-unit change in the corresponding independent variable, holding all other variables constant. The independent variables are:
-   $X_{age}$ : Age of the individual.
-   $X_{degree-score}$ : A measure of the individual's level of education or degree.
-   $X_{finance-score}$ : A measure of the individual's financial stability or status.
-   $X_{health-score}$ : A measure of the individual's overall health.
-   $X_{divorce-score}$ : A measure of the individual's likelihood of divorce or past divorces.
-   $X_{babies-number}$ : The number of babies/children the individual has.

@@

In the Table 3 ,which presented in the table indicate that the sex distribution varies depending on the number of babies in the family.Across all groups, there is a consistent pattern of a higher percentage of females compared to males. The lowest difference in sex distribution is observed among families with no babies, where the female-to-male ratio is relatively closer (54.53% female vs. 45.47% male). As the number of babies in a family increases, the percentage of females also increases, with the highest difference observed in families with 1 and 4 babies. In these groups, the sex distribution is more skewed towards females, with 60.92% and 60.98% female representation, respectively.

```{r}
#| message: false
#| echo: false
#| warning: false

sex_table <- data %>% 
  group_by(babies, sex) %>% 
  summarize(response = n(), variance = var(as.numeric(sex))) %>%
  mutate(percent = round(response/sum(response)*100, 2)) %>%
  select(babies, sex, percent) %>%
  spread(sex, percent)

kable(sex_table, caption = "Relationship between Number of Children and Gender") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)

@@

In the Table 4, which suggests that families with 4 babies have the highest proportion of "happy" individuals at 63.41%. In contrast, families with 3 babies have the highest proportion of "not too happy" individuals at 16.57% and the lowest proportion of "happy" individuals at 50.30%. Families with 0, 1, and 2 babies have relatively similar distributions of happiness levels. It is worth noting that families with 3 babies have the highest proportion of "very happy" individuals at 33.14%, followed closely by families with 2 babies at 32.95%. Families with 4 babies have the lowest proportion of "very happy" individuals at 24.39%.

```{r}
#| message: false
#| echo: false
#| warning: false    
happy_table <- data %>% 
  group_by(babies, happy) %>% 
  summarize(response = n(), variance = var(as.numeric(happy))) %>%
  mutate(percent = round(response/sum(response)*100, 2)) %>%
  select(babies, happy, percent) %>%
  spread(happy, percent)

kable(happy_table, caption = "Happiness Levels Across Families with number of Babies") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)

@@

In Table 5, which appears suggests that individuals without any babies tend to have higher percentages in all health categories, especially in the "poor" health category (93.52%). As the number of babies in a family increases, the proportion of individuals in the "excellent," "fair," and "good" health categories generally decreases. The "poor" health category also follows this trend, but the decrease is less pronounced. Individuals with 1 or 2 babies have relatively similar health status distributions, with slightly higher percentages in the "excellent" and "good" health categories compared to the "fair" and "poor" health categories. Individuals with 3 or 4 babies have the lowest percentages across all health categories.

```{r}
#| message: false
#| echo: false
#| warning: false    
health_happy_table <- data %>%
  group_by(health, babies) %>%
  summarize(response = n(), variance = var(as.numeric(babies))) %>%
  mutate(percent = round(response / sum(response) * 100, 2)) %>%
  select(health, babies, percent) %>%
  spread(babies, percent)

kable(health_happy_table, caption = "Health Status with number of Babies") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)

@@

In Table 6, which appears suggests that families with 0 babies have the highest percentages of financial satisfaction across all categories. As the number of babies in a family increases, the proportion of families in each financial satisfaction category generally decreases. Families with 4 babies have the lowest percentages in all financial satisfaction categories.Notably, the "pretty well satisfied" category shows the steepest decline in percentages as the number of babies increases.Families with 0 babies in this category have the highest percentage of financial satisfaction at 85.47%, while families with 4 babies have the lowest percentage at 0.05%.

```{r}
#| message: false
#| echo: false
#| warning: false    
finance_happy_table <- data %>%
  group_by(finance_satified, babies) %>%
  summarize(response = n(), variance = var(as.numeric(babies))) %>%
  mutate(percent = round(response / sum(response) * 100, 2)) %>%
  select(finance_satified, babies, percent) %>%
  spread(babies, percent)

kable(finance_happy_table, caption = "Table of Financial Satisfaction and Number of babies") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)

@@

@@

Understanding the historical context is also essential in developing interventions that can address the unique challenges faced by individuals during a particular period in time [@h]. For example, the economic conditions during the early 1970s, such as high inflation and high unemployment rates, could have affected individuals' financial stability and, ultimately, their well-being. Interventions that address these economic challenges may have had a significant impact on individuals' well-being during this period.

Furthermore, the study's findings can be compared to more recent data to examine changes in the relationships between different factors and well-being over time[@i]. For example, the impact of divorce on well-being may have changed over time, as societal attitudes towards divorce have shifted. Comparing the study's findings to more recent data can provide insights into how societal changes have influenced individuals' well-being over time.Overall, the historical context is essential in understanding individuals' well-being during a particular period in time. The study's focus on the early 1970s in the United States provides valuable insights into the factors that influenced individuals' well-being during that period. These findings can be used to develop interventions that address the unique challenges faced by individuals during a particular historical context, as well as to compare changes in the relationships between different factors and well-being over time.

@@

The data from [@fig-three] and [@fig-four] further supports the relationship between family size, happiness, and health. The analysis of the number of children had each year from 1972 to 2012 in relation to the level of happiness shows fluctuations in the proportion of individuals who reported feeling "not too happy" and "happy" over the years. This finding highlights the potential influence of sociocultural and economic factors on happiness and family size over time.

>>>>

---
title: "Examining the Multifaceted Impact of Family Size on Happiness, Health, Gender Dynamics, and Financial Well-being"
subtitle: "A Cross-Sectional Cumulative Analysis of U.S. Microdata from 1972-2012"
author: 
  - SHAOHAN CHANG
thanks: "Code and data are available at: https://github.com/lucas11333/Of_Wellbeing"
date: "`r Sys.time()`"
date-format: "D MMMM YYYY"
abstract: "This study explores the multifaceted impact of family size on happiness, health, gender dynamics, and financial well-being by analyzing cross-sectional cumulative microdata from the United States between 1972 and 2012. This study analyzed the relationship between family size, measured by the number of children, and various aspects of family life, such as happiness, health, and financial satisfaction. The findings suggest that happiness levels are highest among childless individuals, while families with one child report the best health status. Additionally, higher financial satisfaction is associated with fewer children in a family. These insights matter as they provide a better understanding of the factors influencing family size and the potential consequences of family planning decisions on overall wellbeing."
format: pdf
number-sections: true
bibliography: references.bib
---

@@

This research investigates the diverse effects of family size on various aspects such as happiness, health, gender roles, and financial stability. The study employs cross-sectional cumulative microdata from the United States spanning from 1972 to 2012.The impact of the number of children in a family on various aspects of family life has been a topic of interest for researchers and policymakers alike [@a]. Family size and structure play a crucial role in shaping the lives of individuals and communities. Understanding the connections between the number of children and factors such as happiness, health, sex distribution, and financial satisfaction can help inform decisions and policies aimed at promoting family well being [@b]. This paper seeks to explore these relationships and shed light on the complex interplay between family size and various aspects of family life.

@@

In @tbl-var contains information on various personal and demographic characteristics of individuals, collected between the years 1972 and 2012. The data-set comprises of 10 variables, namely age, sex, number of babies, divorce status, self-reported health status, financial satisfaction, level of happiness, and additional variables. The data-set contains 38,960 rows, providing ample opportunities to analyze the relationships between these variables and gain valuable insights into the factors that could potentially impact an individual's overall life satisfaction.

@@

In [@fig-one],which analysis explores the relationship between the number of children in a family and happiness levels, categorized as 'very happy,' 'pretty happy,' and 'not too happy.' Although childless families show the highest numbers in each happiness category, the data collected for such families is more abundant, necessitating an analysis of happiness proportions within each family size. The findings reveal no linear correlation between the number of children and overall family happiness, with happiness varying across family sizes without a consistent pattern. The non-uniform distribution of happiness categories among families with different numbers of children implies that factors beyond the number of children also contribute to happiness levels. In conclusion, the data shows variations in happiness levels among families of different sizes but does not establish a definitive correlation between the number of children and overall family happiness. Further research is needed to better understand the complex factors influencing happiness within families.

@@

In [@fig-four], which observed that when individuals have one child, there is a significantly higher proportion of participants who rate their health as 'good' or 'excellent' compared to other categories. In contrast, for participants who do not have any children, the proportion of those who rate their health as 'poor' is the highest, but no clear conclusion can be drawn from this finding. For individuals with three children, the feedback on their health status is evenly distributed among 'good', 'fair', and 'excellent', with a very small proportion of participants reporting 'poor' health. Overall, having one child is associated with the highest proportion of participants reporting 'excellent' health status.

@@

$$
\hat{Y}=\hat{\beta_0} + \hat{\beta_1}X_{finance-score} + \hat{\beta_2}X_{health-score} + \hat{\beta_3}X_{happy} + 
\hat{\beta_6}X_{babies-number} 
$$ {#eq-model}

@@

-   $\hat{Y}$ is predicted value of the dependent variable. This is the outcome we're trying to predict or estimate based on the independent variables.
-   $\hat{\beta_0}$ : The intercept term, which represents the expected value of $\hat{Y}$ when all independent variables are equal to 0.
-   $\hat{\beta_1}$ , $\hat{\beta_2}$ , $\hat{\beta_3}$ are the estimated coefficients for each independent variable. These coefficients indicate the average change in the dependent variable $\hat{Y}$ associated with a one-unit change in the corresponding independent variable, holding all other variables constant. The independent variables are:
-   $X_{age}$ : Age of the individual.
-   $X_{degree-score}$ : A measure of the individual's level of education or degree.
-   $X_{finance-score}$ : A measure of the individual's financial stability or status.
-   $X_{health-score}$ : A measure of the individual's overall health.
-   $X_{divorce-score}$ : A measure of the individual's likelihood of divorce or past divorces.
-   $X_{babies-number}$ : The number of babies/children the individual has.

@@

In the @tbl-sex, which presented in the table indicate that the sex distribution varies depending on the number of babies in the family.Across all groups, there is a consistent pattern of a higher percentage of females compared to males. The lowest difference in sex distribution is observed among families with no babies, where the female-to-male ratio is relatively closer 54.53% female vs. 45.47% male. As the number of babies in a family increases, the percentage of females also increases, with the highest difference observed in families with 1 and 4 babies. In these groups, the sex distribution is more skewed towards females, with 60.92% and 60.98% female representation, respectively.

```{r}
#| label: tbl-sex
#| tbl-cap: Relationship between Number of Children and Gender
#| echo: false
#| warning: false
sex_table <- data %>% 
  group_by(babies, sex) %>% 
  summarize(response = n(), variance = var(as.numeric(sex))) %>%
  mutate(percent = round(response/sum(response)*100, 2)) %>%
  select(babies, sex, percent) %>%
  spread(sex, percent)

kable(sex_table  ) %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)

@@

In the @tbl-happ, which suggests that families with 4 babies have the highest proportion of "happy" individuals at 63.41%. In contrast, families with 3 babies have the highest proportion of "not too happy" individuals at 16.57% and the lowest proportion of "happy" individuals at 50.30%. Families with 0, 1, and 2 babies have relatively similar distributions of happiness levels. It is worth noting that families with 3 babies have the highest proportion of "very happy" individuals at 33.14%, followed closely by families with 2 babies at 32.95%. Families with 4 babies have the lowest proportion of "very happy" individuals at 24.39%.

```{r}
#| label: tbl-happ
#| tbl-cap: Happiness Levels Across Families with number of Babies
#| message: false
#| echo: false
#| warning: false    
happy_table <- data %>% 
  group_by(babies, happy) %>% 
  summarize(response = n(), variance = var(as.numeric(happy))) %>%
  mutate(percent = round(response/sum(response)*100, 2)) %>%
  select(babies, happy, percent) %>%
  spread(happy, percent)

kable(happy_table) %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)

@@

In @tbl-heal, which appears suggests that individuals without any babies tend to have higher percentages in all health categories, especially in the "poor" health category (93.52%). As the number of babies in a family increases, the proportion of individuals in the "excellent," "fair," and "good" health categories generally decreases. The "poor" health category also follows this trend, but the decrease is less pronounced. Individuals with 1 or 2 babies have relatively similar health status distributions, with slightly higher percentages in the "excellent" and "good" health categories compared to the "fair" and "poor" health categories. Individuals with 3 or 4 babies have the lowest percentages across all health categories.

```{r}
#| label: tbl-heal
#| tbl-cap: Health Status with number of Babies
#| message: false
#| echo: false
#| warning: false    
health_happy_table <- data %>%
  group_by(health, babies) %>%
  summarize(response = n(), variance = var(as.numeric(babies))) %>%
  mutate(percent = round(response / sum(response) * 100, 2)) %>%
  select(health, babies, percent) %>%
  spread(babies, percent)

kable(health_happy_table) %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)

@@

In @tbl-fin, which appears suggests that families with 0 babies have the highest percentages of financial satisfaction across all categories. As the number of babies in a family increases, the proportion of families in each financial satisfaction category generally decreases. Families with 4 babies have the lowest percentages in all financial satisfaction categories.Notably, the "pretty well satisfied" category shows the steepest decline in percentages as the number of babies increases.Families with 0 babies in this category have the highest percentage of financial satisfaction at 85.47%, while families with 4 babies have the lowest percentage at 0.05%.

```{r}
#| label: tbl-fin
#| tbl-cap: Table of Financial Satisfaction and Number of babies
#| message: false
#| echo: false
#| warning: false    
finance_happy_table <- data %>%
  group_by(finance_satified, babies) %>%
  summarize(response = n(), variance = var(as.numeric(babies))) %>%
  mutate(percent = round(response / sum(response) * 100, 2)) %>%
  select(finance_satified, babies, percent) %>%
  spread(babies, percent)

kable(finance_happy_table) %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)

@@

## Role of gender distribution in family size and well-being

The observation that families with more children have a higher percentage of females, particularly in families with one and four children, highlights the importance of examining the impact of gender distribution within families.  In many societies, traditional gender roles dictate that women assume primary responsibilities for children, which could lead to a disproportionate burden on women in larger families. This unequal distribution of responsibilities might impact women's well-being, including their mental and physical health, career opportunities, and financial stability [@j]. Moreover, the unequal distribution of resources within families due to gender biases could exacerbate these disparities, further impacting the well-being of women in larger families [@k]. Additionally, societal expectations and norms surrounding gender roles may influence the decision-making process of couples when considering family size. For instance, cultural pressure to have more children or adhere to specific gender roles within a family might contribute to stress, relationship strain, and decreased well-being among family members [@l].

To gain a more comprehensive understanding of the role of gender in family well-being, future research should examine how gender roles and expectations evolve as family size changes, as well as how these dynamics impact overall happiness, health, and financial satisfaction. This research could help inform policies and interventions that promote gender equity within families, ensuring that the well-being of all family members is considered and supported, regardless of family size.

@@

Understanding the historical context is also essential in developing interventions that can address the unique challenges faced by individuals during a particular period in time [@h]. For example, the economic conditions during the early 1970s, such as high inflation and high unemployment rates, could have affected individuals' financial stability and, ultimately, their well-being. Interventions that address these economic challenges may have had a significant impact on individuals' well-being during this period[@i].

@@

The data from [@fig-three] and [@fig-four] further supports the relationship between family size, happiness, and health. The analysis of the number of children had each year from 1972 to 2012 in relation to the level of happiness shows fluctuations in the proportion of individuals who reported feeling "not too happy" and "happy" over the years. This finding highlights the potential influence of horticultural and economic factors on happiness and family size over time.
